---
title: "NY 11 Demographics"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(DT)
library(RColorBrewer)
opts_chunk$set(echo = FALSE, cache = TRUE, cache.path = here("cache/2016_cong/"))
```

```{r process data, include = FALSE}
# read in cd 11 demo shape file
cd11_ed_sf <- read_sf(here("output/geo/cd11_demo_2016.geojson"))

# read in pres results 2016
pres_2016 <- read_csv(
  here("output/elect_results/2016_pres.csv"),
  col_types = cols(year = "i", total_votes = "i", .default = "c")) %>%
  mutate_at(vars(ends_with("_prop")), as.double) %>%
  mutate_at(vars(ends_with("_tot")), as.integer)

# create dem/rep margins, join with ed geo file
pres_2016_margin <- pres_2016 %>%
  mutate(margin = clinton_prop - trump_prop) %>%
  select(elect_dist, total_votes, margin, clinton_tot,
         trump_tot, clinton_prop, trump_prop)

pres_2016_to_map <- left_join(cd11_ed_sf, pres_2016_margin) %>%
  mutate(elect_dist = paste("ED", elect_dist))

pres_2016_close_rep <- pres_2016_to_map %>%
  filter(margin >= -0.1 & margin < 0)

pres_2016_close_dem <- pres_2016_to_map %>%
  filter(margin >= 0 & margin <= 0.1)

# define a little helper function to format percentages
make_pct <- function(x, digits = 1) {
  paste0(formatC(x * 100, digits = digits, format = "f"), "%")
}

# define a little helper function to format $$
make_dollar <- function(x, digits = 0) {
  paste0("$", formatC(x, digits = digits, format = "f", big.mark = ","))
}

# define a little helper function to format margins
make_margin <- function(x, digits = 1) {
  if_else(x > 0,
          paste0("+",formatC(x * 100, digits = digits, format = "f")),
          formatC(x * 100, digits = digits, format = "f")
          )
} 
```

### American Community Survey sociodemographic data at the election district-level

<span style="background-color:lightgreen">Green outlines</span> indicate election districts that lean Republican (between -10 and 0 point margin)  
<span style="background-color:orange">Orange outlines</span> indicate election districts that lean Democratic (between 0 and +10 point margin)  
Click an ED for details  

## Median Household Income
```{r income map, cache = FALSE, message = FALSE, out.width = "100%", out.height = "600px"}

# make a map of reichard/donovan margin 2016, highlight ed +-10pt
tmap_mode("view")
tm_shape(pres_2016_to_map, name = "HH Income") +
  tm_fill(
    col = "med_hhinc",
    palette = "Purples",
    style = "cont",
    contrast = c(0.2, 1),
    title = "Median HH Income",
    textNA = "No Data",
    popup.vars = c(
      "Median HH Income" = "med_hhinc",
      "Dem/Rep Margin" = "margin",
      "Total Votes" = "total_votes"
    ),
    id = "elect_dist",
    popup.format = list(
      med_hhinc = list(fun = make_dollar),
      margin = list(fun = make_margin),
      total_votes = list(format = "f", digits = 0)
      )
  ) +
  tm_borders(col = "darkgray") +
  tm_shape(pres_2016_close_dem, name = "Lean Dem") +
  tm_borders(col = "orange", lwd = 2) +
  tm_shape(pres_2016_close_rep, name = "Lean Rep") +
  tm_borders(col = "green", lwd = 2) +
  tm_view(
    alpha = 0.85,
    basemaps = "Stamen.TonerLite",
    legend.position = c("right", "bottom")
    )
```
<br>
<br>

## Percentage White
```{r race map, cache = FALSE, message = FALSE, out.width = "100%", out.height = "600px"}

# make a map of reichard/donovan margin 2016, highlight ed +-10pt
tmap_mode("view")
tm_shape(pres_2016_to_map, name = "% White") +
  tm_fill(
    col = "pop_white_adj_pct",
    palette = "-RdPu",
    style = "cont",
    contrast = c(0, 0.8),
    title = "% White",
    textNA = "No Data",
    popup.vars = c(
      "% White" = "pop_white_adj_pct",
      "Dem/Rep Margin" = "margin",
      "Total Votes" = "total_votes"
    ),
    id = "elect_dist",
    popup.format = list(
      pop_white_adj_pct = list(fun = make_pct),
      margin = list(fun = make_margin),
      total_votes = list(format = "f", digits = 0)
      )
  ) +
  tm_borders(col = "darkgray") +
  tm_shape(pres_2016_close_dem, name = "Lean Dem") +
  tm_borders(col = "orange", lwd = 2) +
  tm_shape(pres_2016_close_rep, name = "Lean Rep") +
  tm_borders(col = "green", lwd = 2) +
  tm_view(
    alpha = 0.85,
    basemaps = "Stamen.TonerLite",
    legend.position = c("right", "bottom")
    )
```
<br>
<br>

## Percentage with Bachelor's Degree
```{r edu map, cache = FALSE, message = FALSE, out.width = "100%", out.height = "600px"}

# make a map of reichard/donovan margin 2016, highlight ed +-10pt
tmap_mode("view")
tm_shape(pres_2016_to_map, name = "% Bachelors") +
  tm_fill(
    col = "pop_ba_pct",
    palette = "Blues",
    style = "cont",
    contrast = c(0, 0.8),
    title = "% Bachelors",
    textNA = "No Data",
    popup.vars = c(
      "% Bachelors" = "pop_ba_pct",
      "Dem/Rep Margin" = "margin",
      "Total Votes" = "total_votes"
    ),
    id = "elect_dist",
    popup.format = list(
      pop_ba_pct = list(fun = make_pct),
      margin = list(fun = make_margin),
      total_votes = list(format = "f", digits = 0)
      )
  ) +
  tm_borders(col = "darkgray") +
  tm_shape(pres_2016_close_dem, name = "Lean Dem") +
  tm_borders(col = "orange", lwd = 2) +
  tm_shape(pres_2016_close_rep, name = "Lean Rep") +
  tm_borders(col = "green", lwd = 2) +
  tm_view(
    alpha = 0.85,
    basemaps = "Stamen.TonerLite",
    legend.position = c("right", "bottom")
    )
```
<br>
<br>

## Percentage Renters
```{r rent map, cache = FALSE, message = FALSE, out.width = "100%", out.height = "600px"}

# make a map of reichard/donovan margin 2016, highlight ed +-10pt
tmap_mode("view")
tm_shape(pres_2016_to_map, name = "% Renters") +
  tm_fill(
    col = "pop_rent_pct",
    palette = "Greens",
    style = "cont",
    contrast = c(0, 0.8),
    title = "% Renters",
    textNA = "No Data",
    popup.vars = c(
      "% Renters" = "pop_rent_pct",
      "Dem/Rep Margin" = "margin",
      "Total Votes" = "total_votes"
    ),
    id = "elect_dist",
    popup.format = list(
      pop_rent_pct = list(fun = make_pct),
      margin = list(fun = make_margin),
      total_votes = list(format = "f", digits = 0)
      )
  ) +
  tm_borders(col = "darkgray") +
  tm_shape(pres_2016_close_dem, name = "Lean Dem") +
  tm_borders(col = "orange", lwd = 2) +
  tm_shape(pres_2016_close_rep, name = "Lean Rep") +
  tm_borders(col = "green", lwd = 2) +
  tm_view(
    alpha = 0.85,
    basemaps = "Stamen.TonerLite",
    legend.position = c("right", "bottom")
    )
```

### Swing District Detail
*EDs between -10 and +10 margin*  
<br>
```{r data_table, cache = FALSE}
to_table <- pres_2016_to_map %>%
  filter(between(margin, -0.1, 0.1)) %>%
  mutate(
    lean = if_else(
      (margin >= -0.1 & margin < 0), "Republican", "Democratic"
      ),
    margin = margin * 100
    ) %>%
  select(elect_dist, margin, lean, total_votes, med_hhinc,
         pop_white_adj_pct, pop_ba_pct, pop_rent_pct,
         clinton_prop, trump_prop) %>%
  arrange(margin) %>%
  st_set_geometry(NULL)

datatable(
  to_table,
  class = "row-border hover",
  rownames = FALSE,
  colnames = c("Election District", "Margin", "Lean", "Total Votes", 
               "Median HH Income", "White %", "Bachelors %", "Renters %",
               "Clinton %", "Trump %"),
  extensions = c("Buttons", "Scroller", "Responsive"),
  options = list(
    dom = "Bfrtip",
    buttons = c("csv", "print"),
    deferRender = TRUE,
    scrollY = "370px",
    scroller = TRUE
    )
  ) %>%
  formatPercentage(6:10, 1) %>%
  formatRound(2, 1) %>%
  formatCurrency(5, digits = 0)
```


